Two Phase Debiased Calibration Simulation

### 2024-04-23 src/explore/20240411-tpdcsim.qmd
Ran several basic simulations. Additional algorithms did not work.

library(dplyr)
library(GECal)
library(doParallel)
library(doRNG)

@@

  # Phase 1 
  # Poisson sampling
  # mean(pi_i) = 0.1
  # ptemp <- pt(-pop_df$X3 - 2, 3)
  # pi_i <- ifelse(ptemp > 0.7, 0.7, ptemp)
  ptemp <- pt(-pop_df$X3, 3)
  pi_i <- ptemp
  del1 <- rbinom(nrow(pop_df), 1, pi_i)
  p1_df <- pop_df %>%
    mutate(pi1 = pi_i) %>%
    rename(pi2 = X4) %>%
    mutate(del1 = del1) %>%
    filter(del1 == 1)

@@

  # Phase 2
  # del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2)
  del2 <- rbinom(nrow(p1_df), 1, p1_df$pi2 / 2)

@@

@@

  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
  dc_w <-  
    GEcalib(cbind(samp_X, gdi),
            d = rep(1, nrow(p2_df)),
            extra$pop_tot,
            entropy = entropy,
            DS = FALSE)

@@

@@

  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))
  dc_w <-  
    GEcalib(cbind(samp_X, gdi),
            d = rep(1, nrow(p2_df)),
            pop_est,
            entropy = entropy,
            DS = FALSE)

@@

@@

mc_res <- foreach(iter=1:B_sims, .packages = c("dplyr", "GECal")) %dorng% {

  samps <- gen_twophase(pop_df, n1_obs, n2_obs)
  p1_df <- samps[[1]]
  p2_df <- samps[[2]]

  # ptemp <- pt(-pop_df$X3 - 2, 3)
  # pop_pi1 <- ifelse(ptemp > 0.7, 0.7, ptemp)
  ptemp <- pt(-pop_df$X3, 3)
  pop_pi1 <- ptemp

  pistar <- pi_star(p1_df, p2_df, nrow(pop_df), pop_df = pop_df)
  tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
  dcpop <- 
    dc_poptots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2), sum(-pop_pi1))
    )
  dcest <-
    dc_esttots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_gdi = sum(-pop_pi1)
    )

  return(
    tibble(Est = c("pistar", "tpreg", "dcpop", "dcest"),
           Theta = c(pistar[[1]], tpreg[[1]], dcpop[[1]], dcest[[1]]),
           Var = c(pistar[[2]], tpreg[[2]], dcpop[[2]], dcest[[2]]))
  )

@@

# Find any failed estimates
which(is.na(mc_res$Theta))
which(is.na(mc_res$Var))

@@

>>>>

library(dplyr)
library(GECal)
library(nleqslv)
library(doParallel)
library(doRNG)
library(CVXR)

# source("GEcalib_code.R")

@@

  # Phase 1 
  # Poisson sampling
  # mean(pi_i) = 0.1
  ptemp <- pt(-pop_df$X3 - 2, 3)
  pi_i <- ifelse(ptemp > 0.7, 0.7, ptemp)
  del1 <- rbinom(nrow(pop_df), 1, pi_i)
  p1_df <- pop_df %>%
    mutate(pi1 = pi_i) %>%
    mutate(del1 = del1) %>%
    filter(del1 == 1)

@@

  # Phase 2
  # pi2_1 <- p1_df$X4
  p2tmp <- pt(-p1_df$X1 + 1, 3)
  pi2_i <- ifelse(p2tmp > 0.9, 0.9, p2tmp)
  del2 <- rbinom(nrow(p1_df), 1, pi2_i)

  p1_df <- mutate(p1_df, pi2 = pi2_i)

@@

solveEL_prim <- function(xmat, d, xtot) {
  w_vec <- Variable(nrow(xmat))
  obj <- Maximize(sum(d * log(w_vec)))
  constr <- list(t(xmat) %*% w_vec - xtot == 0)
  prob <- Problem(obj, constr)
  prim <- solve(prob)
  resp <- prim$getValue(w_vec)
  return(drop(resp))
}

# resw = resp
# gbar = sum(-log(resw), na.rm = TRUE)
# toremove <- which(is.na(log(resw)))
# zbar = colSums(xmat[-toremove, ])
# zbar = colSums(xmat)
# lam = drop((gbar %*% zbar) %*% MASS::ginv(t(t(zbar)) %*% t(zbar)))
# 
# f(lam, xmat, d, xtot)
# f(res$x, xmat, d, xtot)
# 
# res$x
# lam
# sum(p2_df$Y * resd) / pop_obs
# sum(p2_df$Y * resp) / pop_obs

solveEL <- function(xmat, d, xtot, acc2 = FALSE) {

  f <- function(lam, xmat, d, xtot, returnw = FALSE) {
    w <- -d / drop(xmat %*% lam)

    if (returnw) {
      return(w)
    } else {
      return(drop(w %*% xmat) - xtot)
    }
  }

  j <- function(lam, xmat, d, xtot) {
    t(xmat) %*% diag(d / drop(xmat %*% lam)^2) %*% xmat
  }

  init <- c(rep(0, ncol(xmat) - 1), 1)
  res <- 
  nleqslv(init, f, jac = j, xmat = xmat, d = d, xtot,
          method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

  resw <- f(res$x, xmat, d, returnw = TRUE)

  if (!(res$termcd %in% c(1))) {

    return(NA)

    w_prim <- solveEL_prim(xmat, d, xtot)
    gbar = sum(-log(w_prim), na.rm = TRUE)
    if (any(is.na(log(w_prim)))) {
      toremove <- which(is.na(log(w_prim)))
      zbar = colSums(xmat[-toremove, ])
    } else {
      zbar = colSums(xmat)
    }
    init = drop((gbar %*% zbar) %*% MASS::ginv(t(t(zbar)) %*% t(zbar)))

    res2 <- 
    nleqslv(init, f, jac = j, xmat = xmat, d = d, xtot,
            method = "Newton", control = list(maxit = 1e5, allowSingular = TRUE))

    if (!(res2$termcd %in% c(1))) {
      return(NA)
    }

    resw <- f(res2$x, xmat, d, returnw = TRUE)

  return(resw)

@@

  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))

  # dc_w <-  
  #   GEcalib(cbind(samp_X, gdi),
  #           # d = rep(1, nrow(p2_df)),
  #           d = 1 / (p2_df$pi2),
  #           extra$pop_tot,
  #           entropy = entropy,
  #           DS = FALSE)

  dc_w_alt <- solveEL(cbind(samp_X, gdi), 1 / p2_df$pi2, extra$pop_tot)
  dc_w <- dc_w_alt

@@

  if ("acc2" %in% names(extra)) {
    acc2 <- extra$acc2
  } else {
    acc2 <- FALSE
  }

@@

  samp_X <- cbind(rep(1, nrow(p2_df)), as.matrix(select(p2_df, X1, X2)))

  dc_w_alt <- solveEL(cbind(samp_X, gdi), 1 / p2_df$pi2, pop_est, acc2)
  dc_w <- dc_w_alt

@@

  # FIXME: Change variance estimation to linearization estimate

@@

mc_res <- foreach(iter=1:B_sims,
                  .packages = c("dplyr", "GECal", "CVXR", "nleqslv")) %dopar% {

  set.seed(iter)
  samps <- gen_twophase(pop_df, n1_obs, n2_obs)
  p1_df <- samps[[1]]
  p2_df <- samps[[2]]

  ptemp <- pt(-pop_df$X3 - 2, 3)
  pop_pi1 <- ifelse(ptemp > 0.7, 0.7, ptemp)

  pistar <- pi_star(p1_df, p2_df, nrow(pop_df), pop_df = pop_df)
  tpreg <- tp_reg(p1_df, p2_df, nrow(pop_df))
  # FIXME: GEcalib is not working. It produces results with very high
  # values of the function.
  dcpop <- 
    dc_poptots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_tot = c(nrow(pop_df), sum(pop_df$X1), sum(pop_df$X2), sum(-pop_pi1))
    )
  dcest <-
    dc_esttots(
      p1_df,
      p2_df,
      nrow(pop_df),
      pop_gdi = sum(-pop_pi1)
    )

  return(
    tibble(Est = c("pistar", "tpreg", "dcpop", "dcest"),
           Theta = c(pistar[[1]], tpreg[[1]], dcpop[[1]], dcest[[1]]),
           Var = c(pistar[[2]], tpreg[[2]], dcpop[[2]], dcest[[2]]),
           Iter = iter)
  )

@@

# Find any failed estimates
which(is.na(mc_res$Theta))
which(is.na(mc_res$Var))
mc_res %>%
  mutate(iter = rep(1:B_sims, each = 4)) %>%
  filter(is.na(Theta))

@@

mc_res %>%
  filter(Est == "dcpop") %>%
  arrange(desc(Var))
